---
title: "compartmental model"
author: "Dan Weinberger"
date: "2/1/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(outbreaks)
library(tidyverse)
library(rstan)
library(gridExtra)
rstan_options (auto_write = TRUE)
options (mc.cores = parallel::detectCores ())
set.seed(3) # for reproductibility
```
 
 Follows case study of https://mc-stan.org/users/documentation/case-studies/boarding_school_case_study.html


## Generate some data
```{r}
source('SimPairs.R')
# #1=uncolonized; 2=kid colonized; 3=adult colonized; 4=both colonized
#can think of each column as a household and each row as a day
 table(HH.States[1,])
 table(HH.States[2,])

 
 
 obs1 <- t(apply(HH.States, 1, function(x) table(x)))
 #obs1 <-obs1[100:130,] #burn 100 days of obs
 obs1 <- obs1[100,] #randomly take observation from day 100
 #obs1plot(obs1[1]) 
 
 
```

What we observe is the colonization state at a particular time slice. But colonized people were colonized at a previous time point, and it is possible that a negative person has already cleared the infection. 
If *both* people infected, it is relative easy, we take date of sample, subtract a number drawn from a distribution of duration to get date of infection and infer which person was the primary infected person, then calculate per-day probability of transmission. If *one* person infected, it is more complicated because the uninfected person could have avoided colonization to that time point OR they could have been colonized and cleared already (we just know the are uncolonized at time); so q needs to depend on both probability of avoiding acquistion from the infected person and probability that they had been colonized and cleared it

## first the easier scenario: both people infected
```{r}

# parameterized by mode (m) and standard deviation (sd):
gen.duration <- function(n1=1000, mode.lower, mode.upper, sd.lower=1, sd.upper=1){
        m1 <- runif(n=n1,mode.lower,mode.upper)
        sd1 <- runif(n=n1,sd.lower,sd.upper) 
        
        ra1 <- ( m1 + sqrt( m1^2 + 4*sd1^2 ) ) / ( 2 * sd1^2 )
        sh1 <- 1 + m1 * ra1
        
        duration1 <- rgamma(n=1000,sh1, ra1) #duration infectiousness
        return(duration1)
}

duration.adult <- 1/gen.duration(mode.lower=1/20, mode.upper=1/15,sd.lower=0.05, sd.upper=0.05)
duration.kid <-  1/gen.duration(mode.lower=1/56, mode.upper=1/28,sd.lower=0.01, sd.upper=0.01)

hist(duration.adult)
hist(duration.kid)

```

from day of sample,if both positive, what is probability that kid was colonized first?
```{r}
mean((duration.kid - duration.adult)> 0)
```


If child colonized, what is probability that adult escaped infection
```{r}
prob.inf.day.kid.adult= 0.02 #daily probability transmit from kid to adult
prob.uninf.day.adult <- 1-prob.inf.day.kid.adult
q.adult <- prob.uninf.day.adult^duration.kid
hist(q.adult)
```

If adult colonized, what is probability that kid escaped infection. Even if the per-day probability of transmission is the same, shorter duration means less likely to transmit.
```{r}
prob.inf.day.adult.kid= 0.02 #daily probability transmit from adult to kid
prob.uninf.day.kid <- 1-prob.inf.day.adult.kid
q.kid <- prob.uninf.day.kid^duration.adult
hist(q.kid)
```

# HERE WE GO!!
## Scenarios
(A) If observe child only colonized, it could be because
1. Adult was index colonization, then cleared it
2. Adult was index, transmitted to child, and adult cleared it prior to sample
3. Child was infected from community and has not (yet) transmitted to parent

(B) If observed adult only colonized, it could be because
1. Child was index colonization then cleared it
2. Child was index, transmitted to adult, and child cleared before sample
3. Adult was infected from community and has not (yet) transmitted to parent

(C) If observed child and adult colonized, it could be because
1. Child was index, then infected adult
2. Adult was index, then infected child
3. Adult and child both infected from community

##ADULT INFECTED FIRST
What is probability that an adult is colonized, transmits to kid and then clears prior to sample?
Set time of infection of adult as t=1, and colonized from t=1 to duration.adult; kid can be infected at any point and is colonized for duration.kid.  

```{r}
#Kid can be infected at any point when adult is colonized (1:duration.adult)
day.kid.infected <- 
        sapply(duration.adult, function(x) sample(1:round(x),size=1  ) )
        
day.kid.cleared <- round(day.kid.infected + round(duration.kid))
day.adult.cleared <- round(duration.adult)
#What is time between when kid infected and when adult clears?

t.both1 <- day.adult.cleared - day.kid.infected 

#How much time is adult colonized, but kid is not
##This can happen if adult is colonized first before kid is colonized, or if kid clear before adult clear

t.adult.only <-  (day.kid.infected - 1) + 
                       (day.adult.cleared > day.kid.cleared)*
                                (day.adult.cleared - day.kid.cleared)

t.kid.only <-  (day.kid.cleared > day.adult.cleared)*
                         (day.kid.cleared - day.adult.cleared) 

#t.a1 is number of days from infection of adult to infection of kid
t.a1 <- (day.kid.infected - 1)  

#t.a2 is number of days from clearance of index kid infection to clearance of adult
t.a2 <- (day.kid.cleared > day.adult.cleared)*
                         (day.kid.cleared - day.adult.cleared)

#When both people infected, what proportion of the time when adult infected is kid also infected?
prop.time.co.infect.adult21 <-  t.adult.only/round(duration.adult)
hist(prop.time.co.infect.adult21)

#When both people infected, what proportion of the time when kid infected is adult also infected?
prop.time.co.infect.kid21 <-  1 - t.kid.only/round(duration.kid)
hist(prop.time.co.infect.kid21)




```

## KID INFECTED FIRST
What is probability that a kid is colonized, transmits to adult and then clears prior to sample?
Set time of infection of kid as t=1, and colonized from t=1 to duration.kid; adult can be infected at any point and is colonized for duration.adult.  

```{r}
#adult can be infected at any point when kid is colonized (1:duration.kid)
day.adult.infected <- 
        sapply(duration.kid, function(x) sample(1:round(x),size=1  ) )
        
day.adult.cleared <- round(day.adult.infected + round(duration.adult))
day.kid.cleared <- round(duration.kid)
#What is time between when adult infected and when kid clears?

t.both2 <- day.kid.cleared - day.adult.infected 

#How much time is kid colonized, but adult is not
##This can happen if kid is colonized first before adult is colonized, or if adult clear before kid clear

t.kid.only <-  (day.adult.infected - 1) + 
                       (day.kid.cleared > day.adult.cleared)*
                                (day.kid.cleared - day.adult.cleared)

t.adult.only <-  (day.adult.cleared > day.kid.cleared)*
                         (day.adult.cleared - day.kid.cleared) 

#t.b1 is number of days from infection of index kid to infection of adult
t.b1 <- (day.adult.infected - 1)

#t.b2 is number of days from clearance of index adult to clearance of kid adult
        
t.b2 <- (day.adult.cleared > day.kid.cleared)*
                         (day.adult.cleared - day.kid.cleared)



#When both people infected, what proportion of the time when kid infected is adult also infected?
prop.time.co.infect.kid12 <-  t.kid.only/round(duration.kid)
hist(prop.time.co.infect.kid12)

#When both people infected, what proportion of the time when adult infected is kid also infected?
prop.time.co.infect.adult12 <-  1 - t.adult.only/round(duration.adult)
hist(prop.time.co.infect.adult12)


```


##SIMPLIFY
We just need these pieces

```{r}
duration.adult <- 1/gen.duration(mode.lower=1/20, mode.upper=1/15,sd.lower=0.05, sd.upper=0.05)

day.kid.infected <- 
        sapply(duration.adult, function(x) sample(1:round(x),size=1  ) )

#t.a1 is number of days from infection of adult to infection of kid
t.a1 <- (day.kid.infected - 1)  

#Same for kids
duration.kid <-  
        1/gen.duration(mode.lower=1/56, mode.upper=1/28,sd.lower=0.01, sd.upper=0.01)

day.adult.infected <- 
        sapply(duration.kid, function(x) sample(1:round(x),size=1  ) )
     
#t.b1 is number of days from infection of index kid to infection of adult
t.b1 <- (day.adult.infected - 1)
```

Probabilities.. we want to estimate q1 and q2
outcome is dichotomous (+/- colonized) for parent and child
```{r}

#probability of adult escaping infection from child up to time t
# aka probability that child alone is colonized up to time t (part 1)
probA3.1 <- q1^(t.b1) * (1-q1)

#Probability of adult escaping infection from child completely
# aka probability that child alone is colonized (part 2)
probA3.2 <- q1^round(duration.kid)

#probability of child escaping infection from adult up to time t 
#aka prob that adult along is colonized  up to time t (part 1)
probB3.1 <- q2^(t.a1)  * (1-q2)

#Probability of child escaping infection completely
#Aka prob that adult alone is colonized (part 2)
probB3.2 <- q2^round(duration.adult)




```


